---
title: "X06 Scraping Schedules"
author: "Kris Walker"
date: "10/30/2019"
output: html_document
---

First we load the required packages or install them if they're not already present.
```{r message=FALSE}
pkgs <- c("rvest", "dplyr", "magrittr", "httr", "stringr")
for (pkg in pkgs){
  if (!require(pkg, character.only = TRUE)){
    install.packages(pkg)
    library(pkg)
  }
}
```

Obtain the URL's of the webpages we are interested in and store them into their own variables.
```{r}
csci_spring_2019_url <- "http://ems.csuchico.edu/APSS/schedule/spr2019/CSCI.shtml"
csci_spring_2020_url <- "http://ems.csuchico.edu/APSS/schedule/spr2020/CSCI.shtml"
math_spring_2019_url <- "http://ems.csuchico.edu/APSS/schedule/spr2019/MATH.shtml"
math_spring_2020_url <- "http://ems.csuchico.edu/APSS/schedule/spr2020/MATH.shtml"
```

The function `read_class_schedule()` takes the url passed as an argument, read in the HTML, selects the necessary
nodes, extracts the necessary elements and text and returns a tibble with the data.
```{r}
read_class_schedule <- function(url){
  webpage <- read_html(url)                      #Use rvest to read in the html file.
  #Select the nodes that we want from the HTML.
  selector <- '//*[contains(concat( " ", @class, " " ), concat( " ", "subj", " " ))] | //*[contains(concat( " ", @class, " " ), concat( " ", "cat_num", " " ))] |
     //*[contains(concat( " ", @class, " " ), concat( " ", "sect", " " ))] | //*[contains(concat( " ", @class, " " ), concat( " ", "title", " " ))] |
     //*[contains(concat( " ", @class, " " ), concat( " ", "instructor", " " ))] | //*[contains(concat( " ", @class, " " ), concat( " ", "tot_enrl", " " ))]'
  my_nodes <- webpage %>% html_nodes(xpath = selector)
  
  #Extract the subject column elements and subject text from the webpage.
  subj_elements <- my_nodes %>% 
    html_nodes(xpath = '//*[contains(concat( " ", @class, " " ), concat( " ", "subj", " " ))]')
  subj_text <- html_text(subj_elements)[-1]
  
  #Extract the cat number column elements and cat number text from the webpage.
  cat_num_elements <- my_nodes %>% 
    html_nodes(xpath = '//*[contains(concat( " ", @class, " " ), concat( " ", "cat_num", " " ))]')
  cat_num_text <- html_text(cat_num_elements)[-1]
  
  #Extract the section number column and section number text from the webpage.
  sect_elements <- my_nodes %>% 
    html_nodes(xpath = '//*[contains(concat( " ", @class, " " ), concat( " ", "sect", " " ))]')
  sect_text <- html_text(sect_elements)[-1]
  
  #Extract the class title  column and class title text from the webpage.
  title_elements <- my_nodes %>% 
    html_nodes(xpath = '//*[contains(concat( " ", @class, " " ), concat( " ", "title", " " ))]')
  title_text <- html_text(title_elements)[-1]
  
  #Extract the instructor column and instructor text from the webpage.
  instr_elements <- my_nodes %>% 
    html_nodes(xpath = '//*[contains(concat( " ", @class, " " ), concat( " ", "Instructor", " " ))]')
  instr_text <- html_text(instr_elements)
  
  #Extract the total enrolled column and total enrolled text from the webpage.
  total_enrl_elements <- my_nodes %>% 
    html_nodes(xpath = '//*[contains(concat( " ", @class, " " ), concat( " ", "enrtot", " " ))]')
  total_enrl_text <- html_text(total_enrl_elements)
  
  #Form the returned tibble using the vectors we've extracted.
  schedule_data <- tibble(
    subject = subj_text,
    cat_number = cat_num_text,
    section = sect_text,
    title = title_text,
    instructor = instr_text,
    total_enrolled = total_enrl_text
  )
}
```

Assign tibbles for each of the four webpages, by calling the read_class_schedule() function which extracts the 
necessary data and creates the tibbles.
```{r}
csci_spring_2019_data <- read_class_schedule(csci_spring_2019_url)
csci_spring_2020_data <- read_class_schedule(csci_spring_2020_url)
math_spring_2019_data <- read_class_schedule(math_spring_2019_url)
math_spring_2020_data <- read_class_schedule(math_spring_2020_url)
```

Add term and year columns to the dataframes to keep track of what class was offered when among other reasons.
```{r}
csci_spring_2019_data <- tibble::add_column(csci_spring_2019_data, term = "spring", year = "2019", .before = "subject")
csci_spring_2020_data <- tibble::add_column(csci_spring_2020_data, term = "spring", year = "2020", .before = "subject")
math_spring_2019_data <- tibble::add_column(math_spring_2019_data, term = "spring", year = "2019", .before = "subject")
math_spring_2020_data <- tibble::add_column(math_spring_2020_data, term = "spring", year = "2020", .before = "subject")
```

Combine them all into one dataframe.
```{r}
all_schedule_data <- bind_rows(csci_spring_2019_data, math_spring_2019_data, csci_spring_2020_data, math_spring_2020_data)
```

Let's take a look at the results.
```{r}
head(all_schedule_data, 10)
tail(all_schedule_data, 10)
```

